Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2014-02-01 paper.Rnw
Condensed a few figures, and other small but significant changes

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               fig.align = 'center', 
               tidy = FALSE, 
               comment = NA, 
               cache = TRUE,
               fig.width = 6, 
               fig.height = 6)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw(base_size = 12))
lm_eqn = function(m) {
  l <- list(a = format(coef(m)[1], digits = 2),
      b = format(abs(coef(m)[2]), digits = 2),
      pearson = format(sqrt(summary(m)$r.squared), digits = 3));
  if (coef(m)[2] >= 0)  {
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)~"="~pearson,l)
  } else {
    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)~"="~pearson,l)    
  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load',fig.show='hide'>>=
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1 & templates.mb < 21),
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
  facet_grid(.~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,geom='point',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases))), linetype='solid', width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_linetype(name="Registration Method") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text = element_text(size = 8))
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
  \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-mean}
  \caption{Dice's similarity score vs. atlas and template library size. Dice's
  similarity score of \mb{} segmentations and \psegs{} for 60 ADNI subjects over a
  range of atlas and template library size, registration algorithm, and label
  fusion method on 10 folds of Monte Carlo cross validation. Error bars indicate
  standard error.} 
  \label{fig:ADNI1-xval-k-mean}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-diff}
    \caption{Increase in similarity score over \multiatlas} 
    \label{fig:ADNI1-xval-k-diff}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-variability}
    \caption{Difference in variability} 
    \label{fig:ADNI1-xval-variability}
  \end{subfigure}
  \caption{\textbf{Whole hippocampus segmentation cross-validation on
  Alzheimer's diseased subjects.}
(\ref{fig:ADNI1-xval-k-diff}) shows increase in mean Dice's Similarity Coefficient of \mb{} over
  \multiatlas{} segmentations vs. atlas and template library size when using the
  \ants{} registration method, and majority-vote label fusion over 10 folds of
  Monte Carlo cross validation. Segmentation similarity is computed against
  \psegs{}. Error bars indicate standard error.
(\ref{fig:ADNI1-xval-variability}) compares the variability in Dice scores of \mb{} vs.
  \multiatlas{} segmentations in the validation. Variance of segmentation
  agreement between \mb{} and \multiatlas{} segmentation is computed for each
  subject across over 10 folds of Monte Carlo cross validation. Shown on
  the y-axis is the p-value resulting from a t-test comparing the distribution of
  variances at each parameter setting (atlas/template library size). Only points
  where \mb{} mean variability is lower than \multiatlas{} are shown. Dashed
  lines indicate a p-value of 0.05 and 0.01.}
\end{figure}

>>>>

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               fig.align = 'center', 
               tidy = FALSE, 
               comment = NA, 
               cache = TRUE,
               fig.width = 6, 
               fig.height = 6)
library(ggplot2)
library(Hmisc)
library(reshape2)
library(ADNIMERGE)
library(plyr)
options(digits=3)
theme_set(theme_bw(base_size = 12))
theme_update(
  legend.key = element_rect(fill='white',colour='white'))
lm_eqn = function(m) {
  l <- list(a = format(coef(m)[1], digits = 2),
      b = format(abs(coef(m)[2]), digits = 2),
      pearson = format(sqrt(summary(m)$r.squared), digits = 3));
  if (coef(m)[2] >= 0)  {
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)~"="~pearson,l)
  } else {
    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)~"="~pearson,l)    
  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load',fig.show='hide',fig.width=10>>=
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1 & templates.mb < 21),
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
  facet_grid(.~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,geom='point',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases))), linetype='solid', width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_linetype(name="Registration Method") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.key.size = unit(0.7, 'lines'),
        legend.position = "bottom", 
        legend.box = "horizontal", 
        legend.text = element_text(size=rel(0.7)),
        axis.text = element_text(size = 8))
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
  \centering
  \begin{subfigure}[t]{\textwidth}
    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-mean}
    \caption{DSC vs. atlas and template library size} 
    \label{fig:ADNI1-xval-k-mean}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-diff}
    \caption{Increase in similarity score over \multiatlas} 
    \label{fig:ADNI1-xval-k-diff}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-variability}
    \caption{Difference in variability} 
    \label{fig:ADNI1-xval-variability}
  \end{subfigure}
  \caption{\textbf{Whole hippocampus segmentation cross-validation on
  Alzheimer's diseased subjects with \psegs{}.}
(\ref{fig:ADNI1-xval-k-mean}) Average DSC score of \mb{} with manual
  segmentations for 60 ADNI subjects taken over 10 folds of
  cross-validation at each parameter setting. Error bars indicate standard
  error. 
(\ref{fig:ADNI1-xval-k-diff}) Increase in DSC of \mb{} over \multiatlas{}
  segmentations. 
(\ref{fig:ADNI1-xval-variability}) shows the significance of t-tests comparing
  the variability in DSC scores of \mb{} and \multiatlas{} across validation
  folds. Only points where \mb{} mean variability is lower than \multiatlas{} are
  shown. Dashed lines indicate p-values of 0.05 and 0.01.}
\end{figure}
